---
title: "SIR Algorithm"
author: "Patrick Glettig & Nicolas Wohlgemuth"
date: "03 March 2018"
output: html_notebook
runtime: shiny
---
```{r, echo = FALSE}
library(reshape2)
library(ggplot2)
```

#Initialize the Parameters
For this task, we first need to define all the necessary parameters, namely the population (pop), the number of infected individuals at the beginning (n_0), the frequency of infection (beta) and the frequency of recovery (gamma).

The population parameter however does not represent an entire population, it is simply the size of the population. To have an entire population of individuals, we need to create a vector with the length of the population:

```{r}
pop <- 5000
n_0 <- 5
beta <- 20
gamma <- 1
s <- rep(0, pop)
```

Next, we need to infect n_0 individuals from the population:
```{r}
for (i in 1:n_0){
  s[sample(pop, size = 1)] <- 1
}
```


#SIR Algorithm
The algorithm takes each infected individual, then, based on the infection and recovery frequency it decides whether this individual will recovery or infect another individual in the population. If it infects another individual, the algorithm checks whether the randomly selected individual is already is susceptible (since it cannot be infected if it is recovered) and only then makes it infected. This process goes on until the entire population is recovered and cannot get infected anymore.

```{r, results="hide"}
iter <- 0
recovered <- c()
infected <- c()
susceptible <- c()
while ((sum(s==2) < pop)&&(iter < 1000)){
  pos_i <- which(s==1)
  recovered <- c(recovered, sum(s==2))#recovered
  infected <- c(infected, sum(s==1))#infected
  susceptible <- c(susceptible, sum(s==0))#susceptible
  for (i in pos_i){
    if (runif(1) < (gamma/(beta+gamma))){
      #recovery
      s[i] <- 2
    } else {
      #infection
      j <- sample(pop,1)
      if (s[j] == 0){
        s[j] <- 1
      }
    }
  }
  iter <- iter+1
}
print(iter)
```

Note that recovered individuals are assigned the value 2, infected 1 and suceptible 0. We have added empty vectors that are filled with the number of recovered, infected and susceptible individuals to reconstruct the evolution of the infection.

# Plot the number of recovered individuals as a function of Beta/Gamma
For the plot, we need a vector that divides the number of recovered individuals by Beta/Gamma. As we have already stored the number of recovered individuals in the loop above, we can do this easily with the following code:
```{r}
my_int <- beta/gamma
Recovered_divided <- c()
for (i in 1:length(recovered)){
  Recovered_divided[i] <- recovered[i]/my_int
}
```


Which renders us the following plots:
```{r, echo = FALSE}
sliderInput("pop", label = "Population Size", min = 1000, max = 8000, value = 5000)
numericInput("beta", label = h3("Beta"), value = 20)
numericInput("gamma", label = h3("Gamma"), value = 1)
numericInput("n_0", label=h3("Infected at the beginning"), value = 5)
```
```{r, echo=FALSE,results="hide"}
s <- rep(0, pop)

#Infect a random sample of the size N_0
for (i in 1:n_0){
  s[sample(pop, size = 1)] <- 1
}

#SIR Algorithm Loop
iter <- 0
recovered <- c()
infected <- c()
susceptible <- c()
while ((sum(s==2) < pop)&&(iter < 1000)){
  pos_i <- which(s==1)
  recovered <- c(recovered, sum(s==2))#recovered
  infected <- c(infected, sum(s==1))#infected
  susceptible <- c(susceptible, sum(s==0))#susceptible
    for (i in pos_i){
    if (runif(1) < (gamma/(beta+gamma))){
      #recovery
      s[i] <- 2
    } else {
      #infection
      j <- sample(pop,1)
      if (s[j] == 0){
        s[j] <- 1
      }
    }
  }
  iter <- iter+1
  print(c(sum(s==0),sum(s==1),sum(s==2)))
}
print(iter)

#Calculate beta/gamma for the task
my_int <- beta/gamma
Recovered_divided <- c()
for (i in 1:length(recovered)){
  Recovered_divided[i] <- recovered[i]/my_int
}
```
```{r, echo=FALSE}
renderPlot({

#Use Qplot for the first plot, recovered individuals as a function of Beta/Gamma
qplot(x= seq_along(Recovered_divided),y = Recovered_divided,geom = "line")+
  labs(title = "Recovered Individuals as a function of Beta/Gamma")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank())
})
```

```{r, echo=FALSE}
renderPlot({
#make a dataframe of results for easier ploting in ggplot
evolution <- data.frame(susceptible,infected, recovered)#add the vectors together
evolution$counter <- seq_along(evolution$susceptible)-1#how many interactions/iterations

evolution <- melt(evolution, id.vars = "counter")#reshape the dataframe from wide to long

#GGPlot that shows the evolution of the population
ggplot(evolution, aes(x = counter, y = value, colour = variable)) +
  geom_line(size = 1)+
  scale_colour_hue(l=30)  + #use darker colors
  labs(title = "Evolution of the Infection", y = "Number of Individuals")+
  theme(axis.title.x=element_blank(),#remove x labels and legend title
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.title = element_blank())
})
```